---
title: "Shiny App - MTCars"
author: "Author: Isaac G. Veras"
date: "2023/10/05"
output: 
  html_document:
    theme: cerulean
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo       = FALSE, 
    message    = FALSE, 
    warning    = FALSE, 
    fig.width  = 8, 
    fig.height = 5
)
```


<div class="well">
## 1. Influence of vehicle weight on fuel consumption:

In this presentation, we explore how the weight of a vehicle, represented as 'weight' (`in thousand pounds`), affects fuel consumption, represented as 'consumption'.

### 1.1. Interactive Chart - Model 1 and 2

### Effects of car weight [wt] on miles per gallon [mpg]

</div>
```
NOTE: We present an interactive graph. (Use the slider to choose the weight!)
```

```{r dir, echo=FALSE}
library(shiny)
shinyApp(
        # Define UI for application to plot
        shinyUI(fluidPage(

                # Application title
                titlePanel(""),

                # Sidebar with a slider input for car weight
                sidebarLayout(
                        sidebarPanel(
                                sliderInput("sliderWT", "What is the weight of the car in 1000 lbs?", 1, 6, value = 3),
                                checkboxInput("showModel1", "Show/Hide Model 1", value = TRUE),
                                checkboxInput("showModel2", "Show/Hide Model 2", value = TRUE)
                        ),

                        # Show a plot of the generated distribution
                        mainPanel(
                                plotOutput("carPlot"),
                                h3("Predicted mpg from Model 1:"),
                                textOutput("pred1"),
                                h3("Predicted mpg from Model 2:"),
                                textOutput("pred2")
                        )
                )
        )),

        # Define server logic required to plot
        shinyServer(function(input, output) {
            mtcars$wtif <- ifelse(mtcars$wt - 3 > 0, mtcars$wt - 3, 0)
            model1      <- lm(mpg ~ wt,        data = mtcars)
            model2      <- lm(mpg ~ wtif + wt, data = mtcars)

            model1pred  <- reactive({
                wtInput <- input$sliderWT
                predict(model1, newdata = data.frame(wt = wtInput))
            })

            model2pred  <- reactive({
                wtInput <- input$sliderWT
                predict(model2, newdata = data.frame(wt = wtInput, wtif = ifelse(wtInput - 3 > 0, wtInput - 3, 0)))
            })

            output$carPlot <- renderPlot({
                wtInput    <- input$sliderWT

                plot(mtcars$wt, mtcars$mpg,
                     xlab = "weight in 1000 lbs",
                     ylab = "miles per gallon",
                     bty  = "n",
                     pch  = 16,
                     xlim = c(1, 6),
                     ylim = c(10, 40)
                )
                if(input$showModel1){
                    abline(
                            model1,
                            col = "goldenrod",
                            lwd = 2
                    )
                }
                if(input$showModel2){
                    model2lines <- predict(model2, newdata = data.frame(wt = 1:6, wtif = ifelse(1:6 - 3 > 0, 1:6 - 3, 0)))
                    lines(1:6, model2lines, col = "darkblue", lwd = 2)
                }

                legend(25, 250, c("Model 1 Prediction", "Model 2 Prediction"),
                       pch=16, col = c("goldenrod", "darkblue"), bty = "n", cex = 1.2)
                points(wtInput, model1pred(), col = "goldenrod", pch = 16 , cex = 2)
                points(wtInput, model2pred(), col = "darkblue" , pch = 16 , cex = 2)
            })

            output$pred1 <- renderText({
                model1pred()
            })

            output$pred2 <- renderText({
                model2pred()
            })
        })
)
```


<div class="well">

## 2. Documentation:

### 2.1. The Model explained

 - By selecting the car weight using the slider, the output generates a prediction of mpg from Model 1 [defined as lm(mpg ~ wt)] and a prediction of mpg from Model 2 [defined as lm(mpg ~ wtif + wt)]. The wtif in Model 2 is defined as [ifelse(wt - 3 > 0, wt - 3, 0)].

 - While in Model 1 only a linear model is used, in Model 2 a hockey-stick calculation is applied where the break-point is defined at 3000 lbs car weight. Therefore, we are able to compare the effect of weight on miles per gallon and the differences in these two models.

### 2.2. Conclusion
In conclusion, although the models differ from each other, car weight and miles per gallon have a negative correlation to each other. That means, the higher the weight [wt] the less miles per gallon [mpg] and vice versa.

**Reproducible Pitch Presentation:** <http://i544c.shinyapps.io/Data_Products_Project3>

**GitHUb Repository:** <https://github.com/i544c/Data_Products_Project3>

</div>